Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2022-09-18 RR_manuscript/Draft_RR.tex

The analytic teams differed in their approach to the acoustic analysis of the speech signal, including choosing specific acoustic measures, the temporal window over which they are taken from and how the measures are transformed.
35\% of models used a measure of duration as the outcome variable, 36\% used fundamental frequency (f0), 12\% used vowel formants, 14\% intensity, and 3\% other measures.

45\% of models obtained acoustic measures at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 43\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 4\% from the whole sentence, and 4\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 52 different measurement specifications.
For example, if we look at those analyses who targeted fundamental frequency (f0), it was operationalized in a myriad of ways including as the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement was taken from the interval of a vowel in the article, the adjective or noun, or in the word interval of article, adjective or noun, or the noun phrase interval or the entire phrase.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

The high variability of acoustic measurements is multiplied by the choices of statistical analysis, including the chosen inferential framework, the type of model, and the model specification, like choice of predictors, interactions and group-level (aka random) effects.

The models included a mean of 2 different predictors (defined as variables or columns in the data table).
This means that, in addition to the critical predictor (typicality of the adjective noun combinations), there was on average at least an additional predictor.
Possible information that was used as predictors included the information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, and speaker gender.

@@

This section deals with the meta-analytic analysis of the results submitted by the teams.
We first report on the between-team variability estimate, which is the focus of this study, followed by the meta-analytic intercept estimate.

@@

As explained in the \textbf{Methods} section, the primary aim of this analysis is to assess the degree of between-team variability.
As a measure of between-team variability, we chose to use the meta-analytic group-level standard deviation \(\sigma_{\alpha_{\text{t}}}\)), i.e.~the standard deviation of the group-level effect of team returned by the meta-analytic model.

According to the meta-analytic model, the group-level standard deviation for team is between 0.03 and 0.07 standard units, at 95\% credibility.
This means that the mean deviations of any individual team from the meta-analytic effect estimate range between ±0.06 to ±0.14 (0.3 * 2, 0.07 * 2) in standard units.

\textbf{Non-preregistered}. In our preregistration we did not take into account the possibility that teams might submit multiple analyses/models.
We have run two further non-preregistered meta-analytic models to account for this added hierarchical layer:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A model with model ID instead of team as the group-level effect (\texttt{(1\textbar{}model\_id)}).
\item
  A model with team and model ID nested within team as group-level effects (\texttt{(1\textbar{}team)\ +\ (1\textbar{}team:model\_id}).
\end{enumerate}

Based on the first non-preregistered model, the group-level standard deviation for model ID is between 0.11 and 0.15 standard units, at 95\% credibility.
This means that the deviations of any individual analysis model from the meta-analytic effect estimate range between ±0.22 to ±0.3 (0.11 * 2, 0.15 * 2) in standard units.

The second non-preregistered model yields a similar model ID estimate to that of the first non-preregistered model.
However, the posterior 95\% CrI for between-team variability in the second non-preregistered model is 0 and 0.05 standard units (\(\beta\) = 0.02, SD = 0.01), corresponding to a mean deviation range of ±0.1 standard units.
One interpretation of these estimates is that the variability of effects is larger within teams than it is across teams.

\begin{tabular}{l|l|r|r}
\hline
outcome & temporal\_window & sd\_min & sd\_max\\
\hline
duration & segment & 33.4 & 51.4\\
\hline
duration & word & 32.6 & 119.8\\
\hline
f0 & segment & 40.2 & 44.6\\
\hline
f0 & word & 4.0 & 47.2\\
\hline
intensity & segment & 3.3 & 7.1\\
\hline
intensity & word & 3.5 & 3.9\\
\hline
\end{tabular}

We illustrate what this means in actual acoustic measures by means of examples.
For each of the submitted analyses that included duration, fundamental frequency or intensity as outcome, we have calculated the standard deviation of those measures in the submitted data table.
We then selected values of standard deviations for the three outcome measurements extracted either from specific segments or words across the data tables submitted by the teams (i.e.~how variable were their raw measures).

Table XXX reports the minimum and maximum SD values taken from a representative range.
If we look at segments (e.g.~measures of individual vowels), duration measurements (as taken by the teams who used duration) have standard deviations that range between 33.4 and 51.4 ms.
Given that the estimated between-model variability in standard units as we obtained from the meta-analytic estimate is between 0.11 and 0.15, we can convert the standard units into milliseconds by multiplying them with the standard deviation values in milliseconds: \(33.4 * 0.11 = 3.8\) ms, \(51.4 * 0.11 = 5.7\) ms; \(33.4 * 0.15 = 5\) ms, \(51.4 * 0.15 = 7.7\) ms.
We can thus calculate 95\% intervals of the estimated between-model variability in milliseconds: 7.4-11.2 ms based on the low CI estimate 0.11, 9.8-15.1 ms based on the high CI estimate 0.15.

\begin{verbatim}
## # A tibble: 6 x 5
##   outcome   temporal_window low      high     unit 
##   <chr>     <chr>           <chr>    <chr>    <chr>
## 1 duration  segment         7.4-11.4 10-15.4  ms   
## 2 duration  word            7.2-26.5 9.8-35.8 ms   
## 3 f0        segment         8.9-9.8  12-13.3  hz   
## 4 f0        word            0.9-10.4 1.2-14.1 hz   
## 5 intensity segment         0.7-1.6  1-2.1    dB   
## 6 intensity word            0.8-0.9  1-1.2    dB
\end{verbatim}

@@

The meta-analytic model estimates the range of probable values of the standardized effect size to be between -0.024 and 0.021 standard units (95\% CrI, mean = -0.001).
In other words, our best guess is that typical and atypical word combinations might not differ in acoustic measures (e.g.~duration, f0 etc,) or, if they do, they do so by a maximum of ±0.02 standard units.

\textbf{Non-preregistered}. As mentioned in the previous section, we have run two further models, one using model ID instead of team and the second using team and model ID nested within team.
In the first model, the meta-analytic intercept estimate is between -0.01 and 0.04 standard units (95\% CrI, mean = 0.015).
This suggests that the acoustic measures of typical word combinations are 0.01 standard units lower to 0.04 standard units higher than the measures of atypical word combinations, at 95\% confidence.
Relative to the preregistered model, this model suggests a somewhat positive effect of typicality (although small negative effects are also possible).
The estimated intercept from the second model (with nested group-level effects) is virtually identical to the first (with only model ID as group-level effect).

(ref:meta\_plot1) illustrates the individual intercepts for critical typicality coefficients across models and teams, sorted in ascending order based on their mean.
Given the nature and wide variety of acoustic operationalizations, there is not always a natural interpretation of the scale, but in most cases a positive effect corresponds to typical word combinations eliciting higher acoustic values (e.g.~longer duration, higher f0, etc.), as mentioned above.
Notably, while the majority of models yields inconclusive results, there are 27 model estimates for which the 95\% credible interval does not contain zero (16\%).

@@

After assessing the variability across teams and models, we now turn toward estimating the impact of a series of predictors on the estimated intercept.
There is a lot of variation, raising the question as to whether we can explain some of this variation or whether it is idiosyncratic \citep{breznau2021observing}?

(ref:meta\_plot1), panel C, displays the coefficients alongside 80\% and 95\% credible intervals for all model predictors.

The model suggests that most team-specific predictors yield very small deviations from the meta-analytical estimate and their 95\% credible intervals include zero, leaving us highly uncertain about their direction.
Neither analysts' prior beliefs in the phenomenon (ADD EST + 95CrI), nor their seniority in terms of years after completing their PhD (ADD EST + 95CrI) seem to compellingly affect model estimates.
Similarly, the evaluation of the quality of the analysis from their peers yielded a rather small effect magnitude, again characterized by large uncertainty (ADD EST + 95CrI).
Interestingly, the model uniqueness, i.e.~how unique the choice and combination of predictors is, affects the analysts estimate, with more unique models producing higher positive estimates.
This suggests that the analysts' results very much depend on their choice of predictors.
Looking at the most important choices during measurement, both the acoustic parameter under investigation (e.g.~duration or f0) and the choice of temporal window affected the results.
(ref:meta\_plot1) displays the posterior estimates for the measurement outcome (i.e.~what acoustic dimension was measured, panel A) and measurement window (i.e.~what is the unit over which was measured, panel B).
If an acoustic dimension related to f0 was measured, estimates are lower than the meta analytic estimate.
If on the other hand, vowel formants were measured, estimates are higher than the meta analytic estimate.
Similarly, if acoustic parameters were measured within individual segments (e.g.~the vowel of the noun), estimates are lower than the meta analytic estimate, and if acoustic parameters were measured across whole phrases (e.g.~``the blue banana''), the estimates are generally higher.
In other words, depending on the choice of measurement, analysts arrived at opposite conclusions about how and if typicality is expressed acoustically.

@@

We gave 30 analyst teams the same speech dataset to answer the same research question: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}.
In order to answer this question, teams had to operationalize latent variables within a multidimensional signal, operationalize and chose appropriate predictors, and construct appropriate statistical models to answer the research question.
This complex process has led to a vast ``garden of forking paths'', i.e.~to a wide range of combinations of possible analytical decisions.
Analyses exhibited at least 47 unique ways of operationalizing the acoustic signal alongside 52 unique ways of constructing their statistical model, amounting to 2444 possible unique combinations.
This is just a bare minimum estimate for the vast multiverse of analytical decisions, ignoring for example detailed ways of extracting the chosen acoustic parameters.

Interestingly, the observed variation in reported effect sizes was not predicted by the analysts' prior expectations about the phenomenon.
In fact, teams on average rated the plausibility of the effect as rather high before receiving access to the data.
The observed variation was neither predicted by the analysts' experience in the field nor by the perceived quality of the analysis as judged by other teams.
Analyses received overall high peer ratings for both the acoustic and the statistical analysis, suggesting that reviewers were generally satisfied with the other teams' approaches.
These findings are very much in line with previous crowd-sourced projects that suggest variation between teams is neither driven by perceived quality of the analysis nor by analysts biases or experience \citep[e.g.,][\citet{breznau2021observing}]{silberzahn2018many}.
Given the mounting evidence, \citet{breznau2021observing} conclude that ``{[}\ldots{]} we are left to believe that idiosyncratic uncertainty is a fundamental feature of the scientific process that is not easily explained by typically observed researcher characteristics or analytic decisions.''
Idiosyncratic variation across researchers might be a fact of life which researchers have to acknowledge and integrate into how they evaluate and present evidence.

While properties of the teams did not seem to systematically affect results, teams' estimates seem to depend on certain measurement choices.
Human speech is a complex multidimensional signal.
Researchers need to make choices about what to measure, how to measure it and which temporal unit to measure it in.
Some outcome choices seem to bias the estimates in our data in one direction while others seem to bias estimates into another.
For example, measurements related to fundamental frequency tended to result in lower estimates while measurements related to vowel formants tended to yield higher estimates.
This asymmetry can have several causes.
First, there could be a true underlying relationship between typicality and the speech signal that manifests itself in some measures but not others and/or manifests itself negatively in one acoustic measure but positively in another.
Second, certain measurement choices might be associated with stronger expectations relative to the research question, which might lead to stronger biases.
Many researchers targeted measures related to voice fundamental frequency (f0) since similar functional relationships like information structure and predictability can be expressed via f0 \citep[e.g.][]{grice2017integrating, turnbull2017role}.
Regardless of its cause, we have to conclude that depending on the choice of how the speech signal is operationalized, researchers might find evidence for or against a theoretically relevant effect.

It particularly struck us that most teams submitted many different models.
Teams submitted up to 16 different models to test for a possible relationship between typicality and the speech signal.
Obviously, the complexity of the speech signal lends itself to multiple approaches, but this plurality of hypothesis tests invites bias and can dramatically increase the rate of falsely claiming the presence of an effect \citep{roettger2019researcher}.
When operating within the frequentist inferential framework, testing the same hypothesis with different dependent variables, increases the Type-I error rate if not corrected for \citep[Tukey 1953,][]{benjamini1995controlling}.
It also invites bias, possibly leading to the selective reporting of those tests that yield a desirable outcome \citep[\citet{john2012measuring}, \citet{simmons2011false}]{kerr1998harking} while null results remain unreported \citep[e.g.,][\citet{rosenthal1979file}]{sterling1959publication}.
Fields such as the speech sciences who generate knowledge based on multidimensional data, should be aware of this flexibility and calibrate their confidence in empirical claims accordingly.

@@

A desired outcome of knowing that different but reasonable measurement choices or statistical approaches might lead to different interpretations of research data is to calibrate our (un)certainty in the strength of the collected evidence and, in turn, communicate that (un)certainty appropriately.
The fact that the choice of measurement, measurement window, and predictor choice affect the answer to the research question further suggests that research assumptions and hypotheses should be formulated in much greater detail, particularly so in regards to how measurement systems (here, the acoustic signal) and underlying conceptual constructs (here, the phonetic expression of typicality) relate to each other.
We should ideally specify the link between conceptual construct and quantitative system---the ``derivation chain'' \citep{dubin1970theory, meehl1990summaries}---prior to data collection and analysis, including defining constructs and their relationship within the quantitative system, specifying auxiliary assumptions and boundary conditions, and defining target measurements, statistical expectations and possible (and impossible) effect magnitudes.
Without well defined derivation chains, we ``are not even wrong'' \citep{scheel2022most} because falsified expectations cannot tell us much about the conceptual constructs they are based on when the relationship between the two is underspecified.

@@

Our study has several limitations that need to be considered when evaluating our results.
First, there is potential for bias.
The coordinating authors have engaged with metascientific research before and have been actively involved in methodological debates about scientific practices including transparency and statistical methods.
We have in the past used the lack of standardized analytical approaches as an argument for proposing behavior and policy changes in the field.
This might have biased our own judgment during the analysis which itself came with many researcher degrees of freedom.
We hope we were able to make these degrees of freedom as well as the timing and reasoning of our analytical choices at least detectable.
However, our results should be independently re-analysed and replicated using a different research question.
Moreover, our sample is an opportunity sample.
We have advertised the project through online platforms which might have led to the exclusion of certain potential researcher groups.
While the number of participating teams is larger than most earlier crowdsourcing projects, it is likely to be too small to estimate for example the impact of certain predictors reliably.
Since predictor levels were not systematically distributed across teams, our estimates are characterized by large uncertainty.
This uncertainty is possibly further inflated by the fact that he research question presented to the teams was vague: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
Interpreting the research question/hypothesis differently in terms of its statistical consequences has recently been shown to explain some variation between analysis teams in many-analysis projects \citep{auspurg2021has}.
However, we consider this very underspecification of research hypotheses in the field of speech science (and beyond, see \citet{scheel2022most}) a common phenomenon.
For example, researchers seem to have not yet agreed on how to acoustically measure cross linguistically common phenomena such as word stress \citep[e.g.][]{gordon2017acoustic}.

\begin{itemize}
\tightlist
\item
  We have not investigated researcher positionality, which is known to be an important factor.
\end{itemize}

Despite these all of these possible sources of bias and uncertainty, our study suggest that at a minimum we can expect the between-team and between-analysis variability that we have observed, which is quite substantial and possibly amplifying biases.

Second, the generalizability of our findings to other disciplines on the one hand and specifically other sub-disciplines of the language sciences is, of course, limited.
We focused on quantitative analyses that require operationalization of a multidimensional signal.
While some of our results can be informative for other disciplines that worke with speech or video signals, they might not be very informative for expected variation between analysts working in qualitative research fields.
This is an important point to make because the cognitive sciences in general, and the language sciences in particular have many research areas that are explored with qualitative methods.
It is conceivable that the discussed issues apply differently or not at all to qualitative data analysis.

>>>>

The analytic teams differed in their approach to the acoustic analysis of the speech signal, including choices related to specific acoustic measures, the temporal window over which they are taken from and how the measures are transformed.
35\% of models used a measure of duration as the outcome variable, 36\% used fundamental frequency (f0), 12\% used vowel formants, 14\% intensity, and 3\% other measures.

45\% of models obtained acoustic measures at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 43\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 4\% from the whole sentence, and 4\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 52 different measurement specifications.
For example, if we consider those analyses who targetfundamental frequency (f0), we find that it is operationalized in a myriad of ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

The large decision space related to how the acoustic signal was measured is further expanded by the choices in the statistical analysis, including the chosen inferential framework, the type of model, and the model specification, including choice of predictors, interactions and group-level effects.

The models included a mean of 2 different predictors (defined as variables or columns in the data table).
This means that, in addition to the critical predictor (typicality of the adjective noun combinations), there was on average at least one additional predictor.
Possible information that was used as predictors included the information structure of the sentence, trial number, semantic dimensions of the referent, part of speech, and speaker gender.

@@

This section deals with the meta-analytic analysis of the results submitted by the teams.
First, we report on the between-team variability estimate, which is the focus of this study, followed by the meta-analytic intercept estimate.

@@

The primary aim of this analysis is to assess the degree of between-team.
As a measure of between-team variability, we chose to use the meta-analytic group-level standard deviation \(\sigma_{\alpha_{\text{t}}}\)).

According to the preregistered meta-analytic model, the group-level standard deviation for team is between 0.03 and 0.07 standard units, at 95\% credibility.
This means that the mean deviations of any individual team from the meta-analytic effect estimate range between ±0.06 to ±0.14 (0.3 * 1.96, 0.07 * 1.96) in standard units at 95\% probability.

\textbf{Non-preregistered}. However, we did not take into account that teams might submit multiple analyses/models in our preregistration which, if unaccounted for, violates the independence assumption.
To account for this added layer of dependency, we have run a model with team and model ID nested within team as group-level effects (\texttt{(1\textbar{}team)\ +\ (1\textbar{}team:model\_id}), which allows us to estimate both the between-team variation along with the between-analysis variation.
This analysis was not preregistered and should thus be interpreted with caution.{[}\^{}Note that before this we have also run a model with just model ID as the only (non-nested) group-level effect. The estimated group-level effect of model ID is identical to that of the nested model, so we will not further discuss it.{]}

The nested model yields a posterior 95\% CrI for between-team variability of 0 to 0.05 standard units (\(\beta\) = 0.02, SD = 0.01), corresponding to a mean deviation range of ±0.1 standard units.
The posterior 95\% CrI for between-analysis variability (nested in teams) is 0.11 to 0.15 standard units (\(\beta\) = 0.132, SD = 0.01).
Taken these two models together, it appears as if the variability of reported effects between any model (within team or across) is substantially larger than variability across teams.
We come back to this important observation later.

@@

After assessing the variation between teams and analyses, we now turn towards the meta-analytic estimate of the effect of typicality on acoustic parameters.
The meta-analytic model estimates the range of probable values of the standardized effect size to be between -0.024 and 0.021 standard units (95\% CrI, mean = -0.001).
In other words, and from a very naive statistical perspective, our best guess is that speakers do not encode typicality in the acoustic signal (e.g.~by duration, f0 etc,) or, if they do, they do so by a maximum of ±0.02 standard units.

\textbf{Non-preregistered}. As mentioned in the previous section, we have run an additional model, using team and model ID nested within team as random effects.
In this non-preregistered model, the meta-analytic intercept estimate is between -0.013 and 0.04 standard units (95\% CrI, mean = 0.014).
This suggests that the acoustic measures of typical word combinations are 0.01 standard units lower to 0.04 standard units higher than the measures of atypical word combinations, at 95\% confidence.
Relative to the preregistered model, this model suggests a somewhat positive effect of typicality (although small negative effects are also possible).

(ref:meta\_plot1) illustrates the individual intercepts for critical typicality coefficients across models and teams, sorted in ascending order based on their mean.
Given the nature and wide variety of acoustic operationalizations, there is not always a natural interpretation of the scale, but in most cases a positive effect corresponds to typical word combinations eliciting higher acoustic values (e.g.~longer duration, higher f0, etc.), as mentioned above.
When looking at the raw estimates and their variance (grey triangles and lines), it is striking how much teams' estimates differed.
Estimates ranged between -0.6544004 to 0.9885856 standard units.

While the majority of model estimates and their uncertainty (after shrinkage, black points and lines) yields inconclusive results, there are 27 model estimates for which the 95\% credible interval does not contain zero (16\%).

@@

After assessing the variability across teams and models, we now turn toward estimating the impact of a series of predictors on the observed variation.
There is a large amount of variation between and within teams, raising the question as to whether we can explain some of this variation or whether it is idiosyncratic \citep{breznau2021observing}?

(ref:meta\_plot1), panel C, displays the coefficients for all predictors alongside their 80\% and 95\% credible intervals.
The model suggests that most team-specific predictors yield very small deviations from the meta-analytic estimate and their 95\% credible intervals include zero, leaving us highly uncertain about their direction.
Neither analysts' prior beliefs in the phenomenon (ADD EST + 95CrI), nor their seniority in terms of years after completing their PhD (ADD EST + 95CrI) seem to compellingly affect model estimates.
Similarly, the evaluation of the quality of the analysis from their peers yielded a rather small effect magnitude, again characterized by large uncertainty (ADD EST + 95CrI).
Interestingly, the model uniqueness, i.e.~how unique the choice and combination of predictors is, affects the analysts estimate, with more unique models producing higher positive estimates (ADD EST + 95CrI).
This suggests that the analysts' results very much depend on their choice of predictors.
Looking at the most important choices during measurement, both the acoustic parameter under investigation (e.g.~f0 or formants) and the choice of temporal window affected the results.
(ref:meta\_plot1) displays the posterior estimates for the measurement outcome (i.e.~what acoustic dimension was measured, panel A) and measurement window (i.e.~what is the unit over which the outcome was measured, panel B).
If an acoustic dimension related to f0 was measured, estimates are lower than the meta analytic estimate.
If on the other hand, vowel formants were measured, estimates are higher than the meta analytic estimate.
Similarly, if acoustic parameters were measured across the entire sentence, estimates are lower than the meta analytic estimate.
In other words, depending on the choice of measurement and the measurment window, analysts arrived at opposite conclusions about how and if typicality is expressed acoustically.

@@

We gave 30 analyst teams the same speech data set to answer the same research question: \emph{Do
speakers acoustically modify utterances to signal atypical word combinations?}.
In order to answer this question, teams had to operationalize latent variables within a multidimensional signal, operationalize and chose appropriate predictors, and construct appropriate statistical models to answer the research question.
This complex process has led to a vast ``garden of forking paths'', i.e.~to a wide range of combinations of possible analytical decisions.
Analyses exhibited at least 47 unique ways of operationalizing the acoustic signal alongside 52 unique ways of constructing their statistical model, amounting to 2444 possible unique combinations.
Note that this is a conservative estimate of the multiverse of analytical decisions for our research question (ignoring further degrees of freedom like e.g.~acoustic parameter extraction, outlier treatment, and transformations).

To gain a better intuition about the amount of variability we will ground the results in actual acoustic measures.
We calculated how much raw data measures for duration, fundamental frequency and intensity varied within teams (by extracting their standard deviation).
Consider them an indication of how much said outcome measures generally vary after being extracted from the signal (see xx for an overview ).
We can now use these value of variation and interpret the meta-analytic estimates relative to it.
For example, for those analyses that analyzed the duration of segments (e.g.~the duration of the stressed vowel in banAna), duration measures exhibit standard deviations that range from 33.4 to 51.4 ms.
These standard deviations allow us to convert the meta analytical estimates into milliseconds by multiplying those values.
The reported effect estimates from teams varied between -0.6544004 and 0.9885856 standard units which corresponds to duration values ranging from -21.856972 to 50.8132988.
If we are more conservative and only convert the meta-analytic estimates of between-analysis variation, we obtain an estimate of between-team variability in milliseconds that ranges from 7.4 to 15.1 (at 95\% credibility).\footnote{minimum SD of duration measures multiplied by lower 95\% CrI of meta analysis: \(33.4 * 0.11 = 3.8\) ms; maximum SD of duration measures multiplied by upper 95\% CrI of of meta analysis: \(51.4 * 0.15 = 7.7\) ms}.

While this might not immediately strike one as a lot of variation, it crosses several theoretically relevant thresholds for perception and articulation:
For example, XXms is considered the just noticeable difference, i.e.~the threshold under which human are claimed not to be able to distinguish them \{REF\}.
Another example is the phenomenon of incomplete neutralization, a widely studies phenomenon, which surrounds duration effects ranging from 7-15ms \citep{nicenboim2018using}.
This particular phenomenon has sparked long-lasting methodological and theoretical debates about the very nature of linguistic representations \citep{port2005against}.\\
The fact that there seems to be idios
Expecting random variation between different analysts of 7 to 15 in one or the other direction is thus concerning and might lead to theoretically completely opposing interpretations.

And this variation seems somewhat random.
For example, the observed variation in reported effect sizes was not predicted by the analysts' prior expectations about the phenomenon.
In fact, teams on average rated the plausibility of the effect as rather high before receiving access to the data.
The observed variation was neither predicted by the analysts' experience in the field nor by the perceived quality of the analysis as judged by other teams.
Analyses received overall high peer ratings for both the acoustic and the statistical analysis, suggesting that reviewers were generally satisfied with the other teams' approaches.
These findings are very much in line with previous crowd-sourced projects that suggest variation between teams is neither driven by perceived quality of the analysis nor by analysts biases or experience \citep[e.g.,][\citet{breznau2021observing}]{silberzahn2018many}.
Following \citet{breznau2021observing}, we are bound to conclude that ``{[}\ldots{]} we are left to believe that idiosyncratic uncertainty is a fundamental feature of the scientific process that is not easily explained by typically observed researcher characteristics or analytic decisions.''{[}p.9{]}.
Idiosyncratic variation across researchers might be a fact of life which researchers have to acknowledge and integrate into how they evaluate and present evidence.

While properties of the teams did not seem to systematically affect results, teams' estimates seem to very much depend on certain measurement choices.
Human speech is a complex multidimensional signal.
Researchers need to make choices about what to measure, how to measure it and which temporal unit to measure it in.
Some outcome choices seem to bias the estimates in our data in one direction while others seem to bias estimates into another.
For example, measurements related to fundamental frequency tended to result in lower estimates while measurements related to vowel formants tended to yield higher estimates.
This asymmetry can have several causes.
First, there could be a true underlying relationship between typicality and the speech signal that manifests itself in some measures but not others and/or manifests itself negatively in one acoustic measure but positively in another.
Since we do not know the truth and our best guess based on the meta analysis suggest no relationship, this remains speculative.
Second and independent of a possible true relationship, certain measurement choices might be associated with stronger expectations relative to the research question, which might lead to stronger researcher biases.
Many researchers targeted measures related to fundamental frequency (f0), since similar functional relationships like information structure and predictability can be expressed by f0 \citep[e.g.][]{grice2017integrating, turnbull2017role}.
Moreover, prior work as actually suggested a relationship between typicality and f0 \citep[e.g.][]{dimitrova2008prosodic, dimitrova2009did}.
Researchers might have heard about those findings, nudging their analysis choices slightly into a particular direction.
Regardless of its cause, we have to conclude that depending on the choice of how the speech signal is operationalized, researchers might find evidence for or against a theoretically relevant effect.
This conclusion is further supported by the fact that between-team variability was lower than between model variability.
This is an important observation when put into context of the fact that most teams submitted many different models.
Teams submitted up to 16 different models to test for a possible relationship between typicality and the speech signal.
The complexity of the speech signal lends itself to multiple approaches, but this plurality of hypothesis tests invites bias and can dramatically increase the rate of falsely claiming the presence of an effect \citep{roettger2019researcher}.
When operating within the frequentist inferential framework, testing the same hypothesis with different dependent variables, increases the Type-I error rate if not corrected for \citep[Tukey 1953,][]{benjamini1995controlling}.
It also invites bias, possibly leading to the selective reporting of those tests that yield a desirable outcome \citep[\citet{john2012measuring}, \citet{simmons2011false}]{kerr1998harking} while null results remain unreported \citep[e.g.,][\citet{rosenthal1979file}]{sterling1959publication}.
Fields such as the speech sciences that advance their theories based on multidimensional data should be aware of this flexibility and calibrate their confidence in empirical claims accordingly.

@@

A desired outcome of knowing that different butc reasonable measurement choices or statistical approaches might lead to different interpretations of research data is to calibrate our (un)certainty in the strength of the collected evidence and, in turn, communicate that (un)certainty appropriately.
The fact that the choice of measurement, measurement window, and predictor choice affect the answer to the research question further suggests that research assumptions and hypotheses should be formulated in much greater detail, particularly so in regards to how measurement systems (here, the acoustic signal) and underlying conceptual constructs (here, the phonetic expression of typicality) relate to each other.
We should ideally specify the link between conceptual construct and quantitative system---the ``derivation chain'' \citep{dubin1970theory, meehl1990summaries}---prior to data collection and analysis, including defining constructs and their relationship within the quantitative system, specifying auxiliary assumptions and boundary conditions, and defining target measurements, statistical expectations and possible (and impossible) effect magnitudes.
Without well defined derivation chains, we ``are not even wrong'' \citep{scheel2022most} because falsified expectations cannot tell us much about the conceptual constructs they are based on when the relationship between the two is underspecified.

@@

Our study has several limitations that need to be considered when evaluating our results.
Importantly, there is potential for bias.
The coordinating authors have engaged with metascientific research before and have been actively involved in methodological debates about scientific practices including transparency and statistical methods.
We have in the past used the lack of standardized analytical approaches as an argument for proposing behavior and policy changes in the field.
This might have biased our own judgment during the analysis which itself came with many researcher degrees of freedom.
We hope we were able to make these degrees of freedom as well as the timing and reasoning of our analytical choices at least detectable.
However, our results should be independently re-analysed and replicated using a different research question.
Moreover, our sample is an opportunity sample.
We have advertised the project through online platforms which might have led to the exclusion of certain potential researcher groups.
While the number of participating teams is larger than most earlier crowdsourcing projects, it is likely to be too small to estimate for example the impact of certain predictors reliably.
Since predictor levels were not systematically distributed across teams, our estimates are characterized by large uncertainty.
This uncertainty is possibly further inflated by the fact that he research question presented to the teams was vague: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
Interpreting the research question/hypothesis differently in terms of its statistical consequences has recently been shown to explain some variation between analysis teams in many-analysis projects \citep{auspurg2021has}.
However, we consider this very underspecification of research hypotheses in the field of speech science (and beyond, see \citet{scheel2022most}) a common phenomenon.
For example, researchers seem to have not yet agreed on how to acoustically measure cross linguistically common phenomena such as word stress \citep[e.g.][]{gordon2017acoustic}.
In light of these possible sources of bias, a field can benefit from explicit positionality statement \citep[e.g.,][]{fox2021open}
Researchers do not analyse data in a vaccuum.
It is important recognize and disclose one's positionality, i.e., a reflection about how educational background, social identity, power, experience and context might influence researchers approaches and interpretations.
Despite these all of these possible sources of bias and uncertainty, our study suggest that at a minimum we can expect the between-team and between-analysis variability that we have observed, which is quite substantial and possibly amplifying biases.
Beyond the potential for bias, our study lacks generalizability.
The generalizability of our findings to other disciplines on the one hand and specifically other sub-disciplines of the language sciences is, of course, limited.
We focused on quantitative analyses that require operationalization of a multidimensional signal.
While some of our results can be informative for other disciplines that work with speech or video signals, they might not be very informative for expected variation between analysts working in qualitative research fields.
This is an important point to make because the cognitive sciences in general, and the language sciences in particular have many research areas that are explored with qualitative methods rather than quantitatively.
It is conceivable that the discussed issues apply differently or not at all to qualitative data analysis.
